Life in the canopy: Community trait assessments reveal substantial functional diversity among fern epiphytes

### 2020-02-19 ms/SI.rmd
Improve table S2

```{r trait_data}
trait_values_footer <- "SLA, specific leaf area. Continuous measurements shown as mean, standard deviation (sample size). 'taxon code' is the name used for that taxon in analyses (see Table S1). For data source, L = lab observation of gametophyte cultured from spore, F = field observation of gametophyte subsequently identified using DNA barcoding, T = morphology inferred based on taxonomy, M = measurement of a field-collected specimen; cited references provided at end of SI. 'NA' indicates missing data."

write_si_table(
  table_id = "trait-values", 
  table = trait_data_for_si,
  main_header = si_table_header,
  table_footer = trait_values_footer
)
```

@@

```{r ancova}
# Format table
ancova_table <-
climate_ancova_results %>% 
  mutate(
    p.value = signif(p.value, 2) %>% format.pval(., digits = 1),
    meansq = round_t(meansq, 2),
    term = case_when(
      term == "habit" ~ "Growth habit",
      term == "el" ~ "Elevation",
      term == "habit:el" ~ "Growth habit × elevation",
      TRUE ~ term
    )
  ) %>%
  select(`Response variable` = variable, `Source of variation` = term, `d.f.` = df, MS = meansq, P = p.value)

ancova_footer <- "d.f., degrees of freedom; MS, mean squares. Response variables include mean, minimum ('min'), maximum ('max'), and standard deviation ('sd') of temperature ('temp', °C) and vapor pressure deficit ('vpd', kPa). 'NA' indicates not applicable."

@@

# Reformat variable names and put in preferred order by joining
linear_fit_table <-
tribble(
  ~var, ~Response,
  "temp_max", "Max. temp.",
  "temp_mean", "Mean temp.",
  "vpd_mean", "Mean VPD",
  "vpd_min", "Min. VPD",
  "temp_sd", "SD temp.",
  "vpd_sd", "SD VPD",
  "ntaxa", "Richness",
  "mpd.obs.z.phy", "MPDphy",
  "mntd.obs.z.phy", "MNTDphy",
  "mpd.obs.z.func", "MPDfunc",
  "mntd.obs.z.func", "MNTDfunc",
  "dissection", "Dissection",
  "length", "Frond length",
  "pinna", "Pinna no.",
  "rhizome", "Rhizome diam.",
  "stipe", "Stipe length",
  "sla", "SLA",
  "width", "Frond width"
) %>%
  left_join(combined_summaries, by = "var") %>%
  select(
    Response,
    Model = model_type,
    AICc,
    f = statistic,
    `Adjusted R2` = adj.r.squared,
    P = p.value,
    `d.f.` = df,
    delta_AICc,
    `Moran's I` = morans_I,
    P_moran = I_pval
  )

linear_fit_footer <- "MPDphy, standard effect size of mean phylogenetic distance; MNTDphy, standard effect size of mean nearest taxon distance; MPDfunc, standard effect size of mean functional distance; MNTDfunc, standard effect size of mean nearest functional distance; VPD, vapor pressure deficit, SD, standard deviation; SLA, specific leaf area; AICc, corrected Akaike Information Criterion; d.f., degrees of freedom; P_moran, P value for the null hypothesis that there is spatial autocorrelation in the residuals as measured with Moran's I. Trait names refer to community-weighted mean values (see Methods). For details on calculations of MPD and MNTD, see Methods. 'NA' indicates not applicable."

@@

```{r phylosig-binary-strict}
# Format table
phylosig_binary_strict_table <-
  phylosig_binary_traits_strict %>%
  mutate(D = round_t(D, 2)) %>%
  mutate(trait = factor(
    trait, 
    levels = c("habit", "gemmae", "glands", "hairs","morphotype"))) %>%
  mutate(trait = lvls_revalue(
    trait,
    new_levels = c("Epiphytic growth", "Gemmae", "Glands", "Hairs", "Cordate morphotype"))) %>%
  arrange(trait) %>%
  select(Trait = trait, `Number of presences` = num_present, `Number of absences` = num_absent, D,
         P_rnd = prob_random, P_bm = prob_brownian)

phylosig_binary_strict_footer <- "P_rnd, probability of obtaining the observed value of D under random trait evolution; P_bm, probability of obtaining the observed value of D under simulation of trait evolution by Brownian motion."

@@

P_rnd, probability of obtaining the observed value of \D under random trait evolution; P_bm, probability of obtaining the observed value of \D under simulation of trait evolution by Brownian motion.

@@

```{r gam-fit}
# Format table
gam_fit_table <-
best_fit_div_models_moran %>%
  # Drop frond length and width (highly correlated with rhizome dia.)
  filter(resp_var %in% c("ntaxa", "mpd.obs.z.phy", "mntd.obs.z.phy", 
                   "mpd.obs.z.func", "mntd.obs.z.func",
                   "dissection", "sla", "stipe", "rhizome", "pinna")) %>%
  mutate(resp_var = factor(
        resp_var, 
        levels = c("ntaxa", "mpd.obs.z.phy", "mntd.obs.z.phy", 
                   "mpd.obs.z.func", "mntd.obs.z.func",
                   "dissection", "sla", "stipe", "rhizome", "pinna"))) %>%
  mutate(
    resp_var = lvls_revalue(
        resp_var, 
        new_levels = c("Richness", "MPDphy", "MNTDphy", 
                       "MPDfunc", "MNTDfunc",
                       "Dissection", "SLA", "Stipe length", "Rhizome diam.", "Pinna no.")) 
  ) %>%
  arrange(resp_var, delta_AICc) %>%
  mutate(
    modname = modname %>%
      str_replace_all("\\+", " \\+ ") %>%
      str_replace_all("\\.by\\.", " x ") %>%
      str_replace_all("habit", "Growth habit") %>%
      str_replace_all("temp_mean", "Mean temp.") %>%
      str_replace_all("temp_sd", "SD temp.") %>%
      str_replace_all("temp_max", "Max. temp.") %>%
      str_replace_all("vpd_mean", "Mean VPD") %>%
      str_replace_all("vpd_sd", "SD VPD") %>%
      str_replace_all("vpd_min", "Min. VPD")
  ) %>% 
  select(
    `Metric` = resp_var,
    `Model` = modname,
    AICc,
    `R2` = r2,
    delta_AICc,
    `Moran's I` = morans_I,
    `P_moran` = morans_I_pval
  )

gam_fit_footer <- "MPDphy, standard effect size of mean phylogenetic distance; MNTDphy, standard effect size of mean nearest taxon distance; MPDfunc, standard effect size of mean functional distance; MNTDfunc, standard effect size of mean nearest functional distance; VPD, vapor pressure deficit, SD, standard deviation; AICc, corrected Akaike Information Criterion; P_moran, P value for the null hypothesis that there is spatial autocorrelation in the residuals as measured with Moran's I. Models within delta AICc 3 of the best-fitting model shown. Trait names refer to community-weighted mean values (see Methods). Frond length and width were highly correlated with rhizome diameter and not shown. For details on calculations of MPD and MNTD, see Methods."

>>>>

```{r trait_data}
trait_values_footer <- "SLA, specific leaf area; 'Taxon code', name used for that taxon in analysis (see Table S1); 'Growth habit', growth habit observed in the field; 'Binary growth habit', binary growth habit used in analysis (see Methods). Continuous measurements shown as mean, standard deviation (sample size). For data source, L = lab observation of gametophyte cultured from spore, F = field observation of gametophyte subsequently identified using DNA barcoding, T = morphology inferred based on taxonomy, M = measurement of a field-collected specimen; cited references provided at end of SI. For frond dissection, 1 = simple, 2 = pinnatifid, 3 = one-pinnate, 4 = one-pinnate-pinnatifid, 5 = two-pinnate, 6 = two-pinnate-pinnatifid, 7 = bipinnatifid_or_tripinnatifid, 8 = three-pinnate, 9 = three-pinnate-pinnatifid. 'NA' indicates missing data."

# Format table: capitalize column names
trait_values_table <-
trait_data_for_si %>% 
  rename_all(str_to_sentence) %>% 
  rename(`SLA (sq m per kg)` = `Sla (sq m per kg)`)

write_si_table(
  table_id = "trait-values", 
  table = trait_values_table,
  main_header = si_table_header,
  table_footer = trait_values_footer
)
```

@@

```{r ancova}
# Format table
ancova_table <-
  tribble(
  ~variable, ~Response,
  "temp_max", "Max. temp.",
  "temp_mean", "Mean temp.",
  "temp_sd", "SD temp.",
  "vpd_min", "Min. VPD",
  "vpd_mean", "Mean VPD",
  "vpd_sd", "SD VPD",
) %>%
  left_join(climate_ancova_results, by = "variable") %>% 
  mutate(
    p.value = signif(p.value, 2) %>% format.pval(., digits = 1),
    meansq = round_t(meansq, 2),
    term = case_when(
      term == "habit" ~ "Growth habit",
      term == "el" ~ "Elevation",
      term == "habit:el" ~ "Growth habit × elevation",
      TRUE ~ term
    )
  ) %>%
  select(Response, `Source of variation` = term, `d.f.` = df, MS = meansq, P = p.value)

ancova_footer <- "d.f., degrees of freedom; MS, mean squares. Response variables include mean, minimum, maximum, and standard deviation of temperature ('temp', °C) and vapor pressure deficit ('VPD', kPa). Vapor pressure deficit was square-root transformed prior to analysis. 'NA' indicates not applicable."

@@

# Reformat variable names and put in preferred order by joining
linear_fit_table <-
tribble(
  ~var, ~Response,
  "temp_max", "Max. temp.",
  "temp_mean", "Mean temp.",
  "temp_sd", "SD temp.",
  "vpd_min", "Min. VPD",
  "vpd_mean", "Mean VPD",
  "vpd_sd", "SD VPD",
  "ntaxa", "Richness",
  "mpd.obs.z.phy", "MPDphy",
  "mntd.obs.z.phy", "MNTDphy",
  "mpd.obs.z.func", "MPDfunc",
  "mntd.obs.z.func", "MNTDfunc",
  "dissection", "Dissection",
  "length", "Frond length",
  "pinna", "Pinna no.",
  "rhizome", "Rhizome diam.",
  "stipe", "Stipe length",
  "sla", "SLA",
  "width", "Frond width"
) %>%
  left_join(combined_summaries, by = "var") %>%
  select(
    Response,
    Model = model_type,
    AICc,
    f = statistic,
    `Adjusted R2` = adj.r.squared,
    P = p.value,
    `d.f.` = df,
    `delta AICc` = delta_AICc,
    `Moran's I` = morans_I,
    `P moran` = I_pval
  )

linear_fit_footer <- "MPDphy, standard effect size of mean phylogenetic distance; MNTDphy, standard effect size of mean nearest taxon distance; MPDfunc, standard effect size of mean functional distance; MNTDfunc, standard effect size of mean nearest functional distance; VPD, vapor pressure deficit; SLA, specific leaf area; AICc, corrected Akaike Information Criterion; d.f., degrees of freedom; 'P moran', P value for the null hypothesis that there is spatial autocorrelation in the residuals as measured with Moran's I. Trait names refer to community-weighted mean values (see Methods). For details on calculations of MPD and MNTD, see Methods. 'NA' indicates not applicable."

@@

```{r phylosig-binary-strict}
# Format table
phylosig_binary_strict_table <-
  phylosig_binary_traits_strict %>%
  mutate(D = round_t(D, 2)) %>%
  mutate(trait = factor(
    trait, 
    levels = c("habit", "gemmae", "glands", "hairs","morphotype"))) %>%
  mutate(trait = lvls_revalue(
    trait,
    new_levels = c("Epiphytic growth", "Gemmae", "Glands", "Hairs", "Cordate morphotype"))) %>%
  arrange(trait) %>%
  select(Trait = trait, 
         `Number of presences` = num_present, 
         `Number of absences` = num_absent, 
         D,
         `P rnd` = prob_random, 
         `P bm` = prob_brownian)

phylosig_binary_strict_footer <- "'P rnd', probability of obtaining the observed value of D under random trait evolution; 'P bm', probability of obtaining the observed value of D under simulation of trait evolution by Brownian motion."

@@

'P rnd', probability of obtaining the observed value of \D under random trait evolution; 'P bm', probability of obtaining the observed value of \D under simulation of trait evolution by Brownian motion.

@@

```{r gam-fit}
# Format table

# First make join table for converting variable names, in order
resp_var_table <-
tribble(
  ~resp_var, ~Response,
  "ntaxa", "Richness",
  "mpd.obs.z.phy", "MPDphy",
  "mntd.obs.z.phy", "MNTDphy",
  "mpd.obs.z.func", "MPDfunc",
  "mntd.obs.z.func", "MNTDfunc",
  "dissection", "Dissection",
  "length", "Frond length",
  "width", "Frond width",
  "pinna", "Pinna no.",
  "rhizome", "Rhizome diam.",
  "stipe", "Stipe length",
  "sla", "SLA"
)

gam_fit_table <-
resp_var_table %>%
  left_join(best_fit_div_models_moran, by = "resp_var") %>%
  mutate(resp_var = fct_relevel(resp_var, resp_var_table$resp_var)) %>%
  arrange(resp_var, delta_AICc) %>%
  mutate(
    modname = modname %>%
      str_replace_all("\\+", " \\+ ") %>%
      str_replace_all("\\.by\\.", " x ") %>%
      str_replace_all("habit", "Growth habit") %>%
      str_replace_all("temp_mean", "Mean temp.") %>%
      str_replace_all("temp_sd", "SD temp.") %>%
      str_replace_all("temp_max", "Max. temp.") %>%
      str_replace_all("vpd_mean", "Mean VPD") %>%
      str_replace_all("vpd_sd", "SD VPD") %>%
      str_replace_all("vpd_min", "Min. VPD")
  ) %>% 
  select(
    Response,
    `Model` = modname,
    AICc,
    `R2` = r2,
    `delta AICc` = delta_AICc,
    `Moran's I` = morans_I,
    `P moran` = morans_I_pval
  )

gam_fit_footer <- "MPDphy, standard effect size of mean phylogenetic distance; MNTDphy, standard effect size of mean nearest taxon distance; MPDfunc, standard effect size of mean functional distance; MNTDfunc, standard effect size of mean nearest functional distance; SLA, specific leaf area; VPD, vapor pressure deficit, AICc, corrected Akaike Information Criterion; 'P moran', P value for the null hypothesis that there is spatial autocorrelation in the residuals as measured with Moran's I. Models within delta AICc 3 of the best-fitting model shown. Trait names refer to community-weighted mean values (see Methods). For details on calculations of MPD and MNTD, see Methods."
